---
title: "Advanced ELAN course"
author: "Niko Partanen"
date: "11/16/2017"
output:
  html_document:
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Slides

The course is split along following slides:

- [Introduction](https://langdoc.github.io/elan_lectures/lecture-1)
- [Getting test corpus into R](https://langdoc.github.io/elan_lectures/lecture-2)
- [How does ELAN work?](https://langdoc.github.io/elan_lectures/lecture-3)
- [More advanced example with emuR, Shiny and PraatScript](https://langdoc.github.io/elan_lectures/lecture-4)
- [Some file manipulation examples](https://langdoc.github.io/elan_lectures/lecture-5)

If you have comments or suggestions, any feedback, my email address is nikotapiopartanen@gmail.com. In principle I'm also in Twitter, @nikopartanen, but I don't understand very well how to use it.

## Preparatory work

Before the course starts it can be a good idea to install some of the following tools:

- [R](https://www.r-project.org/)
- [RStudio](https://www.rstudio.com/)
- [Python](https://www.python.org/downloads/)

## Getting course materials

- Download or clone the lecture Rmd files from [here](https://github.com/langdoc/elan_lectures)
- Download or clone the demo corpus from [here](https://github.com/langdoc/testcorpus)

You can do from command line:

    git clone https://github.com/langdoc/elan_lectures
    git clone https://github.com/langdoc/testcorpus

Course materials, or related materials, are also slowly being collected into accompanying [website](https://nikopartanen.github.io/adv_elan_draft/), but this doesn't usually reflect the actual state of things I want to discuss. This work in on draft level, but the idea is to have there one method of discussing things further.

## R packages

You need (at least) following R packages.

- [tidyverse](https://www.tidyverse.org/)
- [shiny]()
- [leaflet]()
- [emuR](https://github.com/IPS-LMU/emuR)
- [meow](https://github.com/achubaty/meow)

## Python

    pip install pympi-ling

## How to load R packages

```{r}
#install.packages("devtools")
#install.packages("tidyverse")

library(devtools)
library(tidyverse)
library(xml2)

#install_github(repo = "langdoc/FRelan")
#install_github(repo = "achubaty/meow")

library(FRelan)
library(meow)

```

## Projects in RStudio

In RStudio you can easily arrange your work around projects. Those can be accessed in right-upper corner of the program. If you work in a project, your working directory is always the one of the project, and you avoid lots of hassle.

In case you need, this is how you can change the working directory.

```{r, eval=FALSE}
getwd()
setwd("~/path/to/some/directory")
```

## Expected file structure

Throughout the course we assume that you have following file structure:

    higher_level_folder
        testcorpus
        elan_lectures
        praat-stuff

## Examples

- Reading ELAN files or individual tiers

```{r}

FRelan::read_eaf(eaf_file = '../testcorpus/kpv_udo20120330SazinaJS-encounter.eaf')
FRelan::read_tier(eaf_file = '../testcorpus/kpv_izva20140330-1-fragment.eaf', linguistic_type = "refT")
FRelan::read_tier(eaf_file = '../testcorpus/kpv_izva20140330-1-fragment.eaf', linguistic_type = "orthT")

```

## Observations

- With Windows backslashes can be strange
- How the folder paths work on Windows?
- Unicode issues: easy to fix with some command, it is about some default settings?
- Some ways of fixing Cyrillic not being displayed properly still resulted some characters not being ok

##  Filenames

Things will generally go better if you never have in filenames:

- Other than ascii characters
- __
- No spaces

## 

```{r, eval=FALSE}

FRelan::read_tier(eaf_file = '../testcorpus/kpv_izva20140330-1-fragment.eaf', linguistic_type = "orthT")

## DOES NOT WORK
## This should be customized somehow
FRelan::read_tier(eaf_file = '../testcorpus/kpv_izva20140330-1-fragment.eaf', linguistic_type = "orthT", 
                  participant_location = "attribute|prefix|suffix")


FRelan::read_eaf('pite.eaf', ind_tier = "reference tier type", sa_tier = "transcription tier type", ss_tier = "token level tier type")
```

## Writing files

In principle you can write something into csv like this.

```{r}

FRelan::read_tier(eaf_file = '../testcorpus/kpv_izva20140330-1-fragment.eaf', linguistic_type = "orthT") %>%
  write_csv('results.csv')

read_csv('results.csv')

```

There are also other ways:

- [googlesheets](https://github.com/jennybc/googlesheets) package allows writing into Google Sheets
- [readxl](http://readxl.tidyverse.org/) should be really good with Excel files

## Improvements

- ind_tier etc to have better names in FRelan functions
- Some stuff there is very messy
- Lots of things are experimental and not supposed to be used by anyone
- Some parts use [XML](https://cran.r-project.org/web/packages/XML/index.html) R package instead of newer [xml2] 

## Custom read_eaf function

This should be adaptable to most of ELAN tier structures, although I admit one maybe has to think a bit what is going on there. In principle one should always be able to recover the logic of ELAN's internal structure like this.

```{r}

read_custom_eaf <- function(path_to_file = '../testcorpus/kpv_udo20120330SazinaJS-encounter.eaf'){

ref <- FRelan::read_tier(eaf_file = path_to_file, linguistic_type = "refT") %>%
  dplyr::select(content, annot_id, participant, time_slot_1, time_slot_2) %>%
  dplyr::rename(ref = content) %>%
  dplyr::rename(ref_id = annot_id)

orth <- FRelan::read_tier(eaf_file = path_to_file, linguistic_type = "orthT") %>%
  dplyr::select(content, annot_id, ref_id, participant) %>%
  dplyr::rename(orth = content) %>%
  dplyr::rename(orth_id = annot_id) # %>%
  # dplyr::rename(ref_id = ref_id) # This is there just as a note

token <- FRelan::read_tier(eaf_file = path_to_file, linguistic_type = "wordT") %>%
  dplyr::select(content, annot_id, ref_id, participant) %>%
  dplyr::rename(token = content) %>%
  dplyr::rename(token_id = annot_id) %>%
  dplyr::rename(orth_id = ref_id)

lemma <- FRelan::read_tier(eaf_file = path_to_file, linguistic_type = "lemmaT") %>%
  dplyr::select(content, annot_id, ref_id, participant) %>%
  dplyr::rename(lemma = content) %>%
  dplyr::rename(lemma_id = annot_id) %>%
  dplyr::rename(token_id = ref_id)

pos <- FRelan::read_tier(eaf_file = path_to_file, linguistic_type = "posT") %>%
  dplyr::select(content, ref_id, participant) %>%
  dplyr::rename(pos = content) %>%
  dplyr::rename(lemma_id = ref_id)

elan <- left_join(ref, orth) %>% 
  left_join(token) %>% 
  left_join(lemma) %>% 
  left_join(pos) %>%
  select(token, lemma, pos, time_slot_1, time_slot_2, everything(), -ends_with('_id'))
  
time_slots <- FRelan::read_timeslots(path_to_file)

corpus <- elan %>% 
  left_join(time_slots %>% rename(time_slot_1 = time_slot_id)) %>%
  rename(time_start = time_value) %>%
  left_join(time_slots %>% rename(time_slot_2 = time_slot_id)) %>%
  rename(time_end = time_value) %>%
  select(token, lemma, pos, participant, time_start, time_end, everything(), -starts_with('time_slot_'))

corpus %>% mutate(filename = path_to_file)

}

```

## Pite case

- Joshua Wilbur has a special case that gave errors, the general start for checking this is below, in principle it can be boiled down into an XPath
- wordT type is incorrectly somewhere where one should have notesT

```{r, eval=FALSE}
elan_file = "file.eaf"
read_xml(elan_file) %>%
  xml_find_all("//TIER[@LINGUISTIC_TYPE_REF='wordT' and starts-with(@TIER_ID, 'notes')]") %>%
    map(~ glue::glue("Fix tier {xml_attr(.x, 'TIER_ID')} in file {elan_file}"))

```
