---
output: 
  revealjs::revealjs_presentation:
    theme: white
    transition: none
    css: custom.css
    self_contained: no
    center: true
    md_extensions: +fenced_code_attributes
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, error=FALSE, dpi = 400,fig.cap = "")

library(tidyverse)
library(xml2)
```

# Getting complicated

## More advanced example

- In this example we are going to use an emuR R package
- Connects to the web services in BAS
- Starting point:
    - Forced alignation tools are good enough
    - We can get from utterance level to phoneme level for free

## {data-background-iframe="http://ips-lmu.github.io/EMU.html"}

## The way it works

- emuR takes specifically formatted files, builds a database out of those, sends something from there to Munchen, does something over there, we get the result back, that can be written into Praat TextGrid
- When we have Praat TextGrids, we can apply PraatScript
- (We can execute the PraatScript safely from Terminal or R)
- You'll see!

##

```{r, echo=FALSE}

read_custom_eaf <- function(path_to_file){
  
  ref <- FRelan::read_tier(eaf_file = path_to_file, linguistic_type = "refT") %>%
    dplyr::select(content, annot_id, participant, time_slot_1, time_slot_2) %>%
    dplyr::rename(ref = content) %>%
    dplyr::rename(ref_id = annot_id)
  
  orth <- FRelan::read_tier(eaf_file = path_to_file, linguistic_type = "orthT") %>%
    dplyr::select(content, annot_id, ref_id, participant) %>%
    dplyr::rename(orth = content) %>%
    dplyr::rename(orth_id = annot_id)
  
  token <- FRelan::read_tier(eaf_file = path_to_file, linguistic_type = "wordT") %>%
    dplyr::select(content, annot_id, ref_id, participant) %>%
    dplyr::rename(token = content) %>%
    dplyr::rename(token_id = annot_id) %>%
    dplyr::rename(orth_id = ref_id)

  lemma <- FRelan::read_tier(eaf_file = path_to_file, linguistic_type = "lemmaT") %>%
    dplyr::select(content, annot_id, ref_id, participant) %>%
    dplyr::rename(lemma = content) %>%
    dplyr::rename(lemma_id = annot_id) %>%
    dplyr::rename(token_id = ref_id)
  
  pos <- FRelan::read_tier(eaf_file = path_to_file, linguistic_type = "posT") %>%
    dplyr::select(content, ref_id, participant) %>%
    dplyr::rename(pos = content) %>%
    dplyr::rename(lemma_id = ref_id)
  
  elan <- left_join(ref, orth) %>% 
    left_join(token) %>% 
    left_join(lemma) %>% 
    left_join(pos) %>%
    select(token, lemma, pos, time_slot_1, time_slot_2, everything(), -ends_with('_id'))
    
  time_slots <- FRelan::read_timeslots(path_to_file)
  
  elan %>% left_join(time_slots %>% rename(time_slot_1 = time_slot_id)) %>%
    rename(time_start = time_value) %>%
    left_join(time_slots %>% rename(time_slot_2 = time_slot_id)) %>%
    rename(time_end = time_value) %>%
    select(token, lemma, pos, participant, time_start, time_end, everything(), -starts_with('time_slot_')) %>%
    mutate(session_name = str_extract(path_to_file, '[^/]+(?=.eaf)')) %>%
    mutate(filename = path_to_file)

  }

```

- It's often good idea to do all modification and variable creation in one spot…

```{r}
eafs <- dir(path = '~/github/test-corpus/', pattern = 'eaf$', full.names = TRUE)

corpus <- eafs %>% map(read_custom_eaf) %>% bind_rows()

corpus <- corpus %>% 
  mutate(time_duration = time_end - time_start) %>%
  mutate(audio_file = str_replace(filename, 'eaf$', 'wav')) %>%
  mutate(orth_trimmed = str_replace_all(orth, c('[:punct:]' = '', 
                                                '\\s+' = ' '))) %>%
  filter(! participant == 'NTP-M-1986') %>% # just getting rid of myself
  select(orth_trimmed, time_start, time_end, time_duration, everything())

```

##

```{r}
plot(density(corpus$time_duration))
```

##

- One type of data emuR can handle is audio file + matching text file
- It is also picky about the audio files

```{r}
library(exifr)
corpus %>% distinct(audio_file) %>%
        pull(audio_file) %>%
        map(~ exifr::read_exif(.x)) %>% bind_rows() %>%
        rename(audio_file = SourceFile) %>%
  select(BitsPerSample, Duration, FileType, NumChannels, everything())
```

- Too many channels! The web service will crash if it sees this!

##

```{r, eval=FALSE}
library(glue)

corpus %>% 
  distinct(audio_file) %>% 
  pull(audio_file) %>% 
  walk(~ {
    seewave::sox(glue("{.x} -c 1 {str_replace(.x, 'wav$', '-mono.wav')}"))
    })
```


##

- Let's define a function that cuts ELAN files by reference
- seewave package can access sox from R

```{r}
cut_elan_ref <- function(audio_file, reference_id, start, duration){
  
  if (dir.exists('~/github/test-corpus/reference_clips') == FALSE) {
    dir.create('~/github/test-corpus/reference_clips')
  }

  seewave::sox(command = glue("{str_replace(audio_file, 'wav$', '-mono.wav')} ~/github/test-corpus/reference_clips/{reference_id}.wav trim {start / 1000} {duration / 1000}"))
  
}
```

##

- walk is same as map, but it doesn't output anything, you use it for side-effects (like writing file)
    - Aaaactually, it silently returns what it got

```{r, eval=FALSE}
corpus %>% distinct(audio_file, ref, time_start, time_duration, orth_trimmed) %>%
  split(.$ref) %>%
  walk(., ~ cut_elan_ref(.x$audio_file, .x$ref, .x$time_start, .x$time_duration)) %>%
  walk(., ~ write_lines(.x$orth_trimmed[1], path = glue::glue('~/github/test-corpus/reference_clips/', .$ref[1], '.txt')))
```

## {data-background-image="https://i.imgur.com/XC7Xo3m.png" background-image-size="65%"}

## 

```{r, eval=FALSE}

library(emuR)
convert_txtCollection(dbName = 'test-corpus', 
                            sourceDir = '~/github/test-corpus/reference_clips', 
                            targetDir = '.', 
                            txtExtension = '.txt', 
                            mediaFileExtension = 'wav', 
                            attributeDefinitionName = 'orth')

dbHandle = load_emuDB('test-corpus_emuDB', verbose = F)
```

##

```{r, eval=FALSE}
runBASwebservice_g2pForTokenization(handle = dbHandle,
  transcriptionAttributeDefinitionName = 'orth', language = 'rus-RU',
  orthoAttributeDefinitionName = 'ORT', resume = FALSE,
  verbose = TRUE)

runBASwebservice_g2pForPronunciation(handle = dbHandle,
                  orthoAttributeDefinitionName = 'ORT',
                  language = 'und', 
                  canoAttributeDefinitionName = 'KAN', 
                  params = list(embed = 'maus', imap=RCurl::fileUpload("~/github/test-corpus/kpv-sampa.txt")), 
                  resume = FALSE, 
                  verbose = TRUE)

runBASwebservice_maus(handle = dbHandle,
                      canoAttributeDefinitionName = 'KAN',
                      language = 'rus-RU',
                      mausAttributeDefinitionName = 'MAUS',
                      chunkLevel = NULL,
                      turnChunkLevelIntoItemLevel = TRUE,
                      perspective = 'default',
                      resume = FALSE,
                      verbose = TRUE)

export_TextGridCollection(dbHandle, targetDir = '~/github/test-corpus/praat', attributeDefinitionNames = c('ORT', 'KAN', 'MAUS'))
```

## {data-background-image="https://i.imgur.com/C7HfnXx.png" background-image-size="65%"}

## {data-background-image="https://i.imgur.com/Mh6ev8o.png" background-image-size="65%"}

##

### This is where one starts to think:

### "should had I checked all transcriptions once more before I did this?"

## {data-background-image="https://i.imgur.com/hxdO7Oy.png" background-image-size="60%"}

# Integrating tools

## Shiny

- Earlier we had an example of a JavaScript based web content
- Nice, but still limited
- It is possible to build small applications also in R
- Can be hosted for free online, but things get fast tricky
- Anyway having something running and up on server is already bit more complex
- Generally very good for prototyping
    - Compact, lots of examples, logic easy to follow
    - Again, probably it all could be done from scratch in JavaScript

## PraatScript

- A scripting language that allows using Praat functionalities
- Works within Praat, and outside Praat
- Somewhat popular: internet is full of examples
- Not entirely easy or intuitive, but not that bad
    - Follows a lot what happens in Praat GUI

##

```{r}

```

