---
output: 
  revealjs::revealjs_presentation:
    theme: white
    transition: none
    css: custom.css
    self_contained: no
    center: true
    md_extensions: +fenced_code_attributes
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, error=FALSE, dpi = 400,fig.cap = "")

library(tidyverse)
library(xml2)
```

# How does ELAN work?

## Basically it …

- Saves things into XML file
- Saves something when you press save
- Finishes some things when exiting program
- pfsx files contain
    - View settings
    - Last opening location
- Does some validation
    - One when opening the file
    - One with Multiple files search
- Opening file in ELAN and touching something has consequence of XML being pretty printed

## Later validation more strickt

## {data-background="https://i.imgur.com/SNqdMH5.png" data-background-size="75%"}

## How is ELAN file structured?

```{r}
read_xml('../adv_elan_draft/notebooks/test.eaf') %>% xml_structure()
```

## What we want from it?

- Participant, session name, tier content, time codes
- Logical combination of the above
- Good test for internal logic of an ELAN tier structure!

## Example file

## {data-background="https://i.imgur.com/CIVMcYg.png" data-background-size="75%"}

## Where are the annotations?

- On individual tiers, for each speaker on their own structure
- The relations are stored within tier ID's
- But not if you use included in as linguistic type!

## {data-background="https://i.imgur.com/mk4o51D.png" data-background-size="75%"}

## {data-background="https://i.imgur.com/oQZh2jI.png" data-background-size="75%"}

## Where are the participants?

## {data-background="https://i.imgur.com/SkV4zJS.png" data-background-size="75%"}

## {data-background="https://i.imgur.com/6L86tFg.png?1" data-background-size="75%"}

## And time codes?

## {data-background="https://imgur.com/WvDRTz6.png" data-background-size="65%"}

## Notice with time codes…

- Time codes are just listed, annotations tell what are starts and ends
- But I think it uses one time slot only once?
- Times are in milliseconds

## What does this mean?

- Parsing ELAN file content demands walking through the logic in tier structure
- At times there is little logic :(
- However …

##

### bad structure
### always wins
### inconsistent structure

# Parsing ELAN file to R

##

- In earlier example we used read_eaf() function
- Nice, but demands quite specific structure
    - Although customizable
- There are also functions read_tier() and read_timeslot()

```{css echo=FALSE}

```


```{r}
library(kableExtra)
library(FRelan)
read_tier(eaf_file = '~/github/test-corpus/kpv_izva20140330-1-fragment.eaf', 
          linguistic_type = 'wordT')

```


##

- The output we get is directly from the ELAN XML file
- Let's look into some other types of tier

```{r}
read_tier(eaf_file = '~/github/test-corpus/kpv_izva20140330-1-fragment.eaf', 
          linguistic_type = 'refT')
```

## What's going on?

- Whatever the content, it gets returned as "content"
- If there are time slot values, those are picked
- Participant and tier id are picked as well
- We also get annotation id and reference id
    - But as we understand, one tiers id is anothers reference

## 

```{r}

path_to_file = '~/github/test-corpus/kpv_udo20120330SazinaJS-encounter.eaf'

ref <- FRelan::read_tier(eaf_file = path_to_file, linguistic_type = "refT") %>%
  dplyr::select(content, annot_id, participant, time_slot_1, time_slot_2) %>%
  dplyr::rename(ref = content) %>%
  dplyr::rename(ref_id = annot_id)

orth <- FRelan::read_tier(eaf_file = path_to_file, linguistic_type = "orthT") %>%
  dplyr::select(content, annot_id, ref_id, participant) %>%
  dplyr::rename(orth = content) %>%
  dplyr::rename(orth_id = annot_id) # %>%
  # dplyr::rename(ref_id = ref_id)

token <- FRelan::read_tier(eaf_file = path_to_file, linguistic_type = "wordT") %>%
  dplyr::select(content, annot_id, ref_id, participant) %>%
  dplyr::rename(token = content) %>%
  dplyr::rename(token_id = annot_id) %>%
  dplyr::rename(orth_id = ref_id)

lemma <- FRelan::read_tier(eaf_file = path_to_file, linguistic_type = "lemmaT") %>%
  dplyr::select(content, annot_id, ref_id, participant) %>%
  dplyr::rename(lemma = content) %>%
  dplyr::rename(lemma_id = annot_id) %>%
  dplyr::rename(token_id = ref_id)

pos <- FRelan::read_tier(eaf_file = path_to_file, linguistic_type = "posT") %>%
  dplyr::select(content, ref_id, participant) %>%
  dplyr::rename(pos = content) %>%
  dplyr::rename(lemma_id = ref_id)

```

##

```{r}
elan <- left_join(ref, orth) %>% 
  left_join(token) %>% 
  left_join(lemma) %>% 
  left_join(pos) %>%
  select(token, lemma, pos, time_slot_1, time_slot_2, everything(), -ends_with('_id'))
  
time_slots <- FRelan::read_timeslots(path_to_file)

corpus <- elan %>% 
  left_join(time_slots %>% rename(time_slot_1 = time_slot_id)) %>%
  rename(time_start = time_value) %>%
  left_join(time_slots %>% rename(time_slot_2 = time_slot_id)) %>%
  rename(time_end = time_value) %>%
  select(token, lemma, pos, participant, time_start, time_end, everything(), -starts_with('time_slot_'))
```

##

```{r}
corpus
```

```
read_custom_eaf <- function(path_to_file){
  all the code from above...
}

read_custom_eaf(path_to_file = 'path/to/my_file.eaf')

```


## What can go wrong?

- Tier doesn't exist
- Tier has a different name
- Tier types are different
- XML is malformed
- …

# Parsing actual corpus

## Looping through files

- In R there are plenty of ways not to write a for-loop
- Idea is always same: take multiple items of something and do something for every unit
- All paths to the files in corpus is a good starting point

```{r, echo=FALSE}
read_custom_eaf <- function(path_to_file){
  
  ref <- FRelan::read_tier(eaf_file = path_to_file, linguistic_type = "refT") %>%
    dplyr::select(content, annot_id, participant, time_slot_1, time_slot_2) %>%
    dplyr::rename(ref = content) %>%
    dplyr::rename(ref_id = annot_id)
  
  orth <- FRelan::read_tier(eaf_file = path_to_file, linguistic_type = "orthT") %>%
    dplyr::select(content, annot_id, ref_id, participant) %>%
    dplyr::rename(orth = content) %>%
    dplyr::rename(orth_id = annot_id)
  
  token <- FRelan::read_tier(eaf_file = path_to_file, linguistic_type = "wordT") %>%
    dplyr::select(content, annot_id, ref_id, participant) %>%
    dplyr::rename(token = content) %>%
    dplyr::rename(token_id = annot_id) %>%
    dplyr::rename(orth_id = ref_id)

  lemma <- FRelan::read_tier(eaf_file = path_to_file, linguistic_type = "lemmaT") %>%
    dplyr::select(content, annot_id, ref_id, participant) %>%
    dplyr::rename(lemma = content) %>%
    dplyr::rename(lemma_id = annot_id) %>%
    dplyr::rename(token_id = ref_id)
  
  pos <- FRelan::read_tier(eaf_file = path_to_file, linguistic_type = "posT") %>%
    dplyr::select(content, ref_id, participant) %>%
    dplyr::rename(pos = content) %>%
    dplyr::rename(lemma_id = ref_id)
  
  elan <- left_join(ref, orth) %>% 
    left_join(token) %>% 
    left_join(lemma) %>% 
    left_join(pos) %>%
    select(token, lemma, pos, time_slot_1, time_slot_2, everything(), -ends_with('_id'))
    
  time_slots <- FRelan::read_timeslots(path_to_file)
  
  elan %>% left_join(time_slots %>% rename(time_slot_1 = time_slot_id)) %>%
    rename(time_start = time_value) %>%
    left_join(time_slots %>% rename(time_slot_2 = time_slot_id)) %>%
    rename(time_end = time_value) %>%
    select(token, lemma, pos, participant, time_start, time_end, everything(), -starts_with('time_slot_')) %>%
    mutate(session_name = str_extract(path_to_file, '[^/]+(?=.eaf)'))

  }

```


```{r}
dir(path = '~/github/test-corpus', pattern = '.eaf$', full.names = TRUE)
```

Or:

```{r}
dir(path = '~/github/test-corpus', pattern = '.+izva.+eaf$', full.names = TRUE)
```

##

```{r}
elan_files <- dir(path = '~/github/test-corpus', pattern = '.eaf$', full.names = TRUE)
elan_files %>% map(read_custom_eaf)
```

- This list format is useful for testing!

##

```{r, echo=FALSE}
library(xml2)
read_cmdi <- function(cmdi_file){ # this defines the function
  read_xml(cmdi_file) %>% # reads the xml
  xml_find_all('//cmd:Actor') %>% # finds all Actor nodes
  map(~ tibble(participant = .x %>% xml_find_first('./cmd:Code') %>% xml_text,
               session_name = .x %>% xml_find_first('../../cmd:Name') %>% xml_text,
               year_birth = .x %>% xml_find_first('./cmd:BirthDate') %>% xml_text,
               year_rec = .x %>% xml_find_first('../../cmd:Date') %>% xml_text,
               role = .x %>% xml_find_first('./cmd:Role') %>% xml_text,
               sex = .x %>% xml_find_first('./cmd:Sex') %>% xml_text,
               session_address = .x %>% xml_find_first('../../cmd:Location/cmd:Address') %>% xml_text,
               session_country = .x %>% xml_find_first('../../cmd:Location/cmd:Country') %>% xml_text,
               session_location = paste0(session_address, ', ', session_country),
               education = .x %>% xml_find_first('./cmd:Education') %>% xml_text,
               name_full = .x %>% xml_find_first('./cmd:FullName') %>% xml_text)) %>% 
    bind_rows() # After everything is collected into tibble/dataframe,
                # we can just bind the rows together
}
```


```{r}
elan_corpus <- elan_files %>% map(read_custom_eaf) %>% bind_rows()
meta <- dir('~/github/test-corpus/', pattern = 'cmdi$', full.names = TRUE) %>% 
  map(read_cmdi) %>% 
  bind_rows()

test_corpus <- left_join(elan_corpus, meta) %>% left_join(read_csv('coordinates.csv'))
```

##

- I can recommend trying View function in RStudio

```{r, eval=FALSE}
write_rds(test_corpus, 'test_corpus.rds')
test_corpus <- read_rds('test_corpus.rds')
source('parse_corpus.R')
test_corpus <- monster_function_that_does_everything()
test_corpus %>% View
```

##

```{r}
test_corpus %>% count(participant)
test_corpus %>% count(session_location)
test_corpus %>% count(year_birth)
```

## Next: **More advanced example** or **Trying this further**?
